---
title: "R Notebook"
output: html_notebook
---

Source: http://markovjumps.blogspot.be/2011/12/r-array-to-rcpparmadillo-cube.html
    https://stackoverflow.com/a/12570932/5270873
    
RcppArmadillo has a proper 'Cube' type generalizing matrices to 3-d.  
RcppArmadillo has `Cube` as a first class object.

```{r}
require(inline)
require(RcppArmadillo)
 
src <- '
    using namespace Rcpp;
     
    NumericVector vecArray(myArray);
        IntegerVector arrayDims = vecArray.attr("dim");
        arma::cube cubeArray(vecArray.begin(), 
                             arrayDims[0], arrayDims[1], arrayDims[2], false);
 
        // change one element in the array/cube
        cubeArray(0,0,0) = 518;  
 
    return(wrap(cubeArray));  
'
```


```{r}
readCube = cxxfunction(signature(myArray="numeric"), 
                       body=src, 
                       plugin="RcppArmadillo")

set.seed(345)
testArray = array(rnorm(18), dim=c(3,3,2))
print(testArray)
```


```{r}
readCube(testArray)
```



```{r}
require(inline)
require(RcppArmadillo)
 
example <- '
    using namespace Rcpp;
    arma::cube x(1,2,3);
    arma::cube y = arma::cube(4,5,6);

    y.ones();
    
    arma::mat B = arma::mat(4,5);
    B.zeros();
    y.slice(2) = B;

    arma::cube p(5,6,7) ;
    p.fill(7);
    arma::cube q = y + y;
    arma::cube r = y %  y; // %	Schur product: element-wise multiplication of two objects

    arma::cube::fixed<4,5,6> f;
    f.ones();

    x.ones();

    p(0,0,0) = 1;

    return(wrap( r ));
'

cubeExample = cxxfunction(body=example, 
                       plugin="RcppArmadillo")


myCube <- cubeExample()
dim(myCube)
myCube
```

## subsube, slicing, save

```{r}
require(inline)
require(RcppArmadillo)
 
subcube <- '
    using namespace Rcpp;
    
    arma::cube A = arma::cube(2,3,4);
    A.zeros();

    arma::mat B = A.slice(1);   // each slice is a matrix
    
    A.slice(0) = arma::mat(2,3);
    A.slice(0)(1,2) = 99.0;

    A.subcube(0,0,1,  1,1,2)   = arma::cube(2,2,2);
    A( arma::span(0,1), arma::span(0,1), arma::span(1,2) ) = arma::cube(2,2,2);
    A( 0,0,1, arma::size(2,2,2) )              = arma::cube(2,2,2);

    // add 123 to all elements of A greater than 0.5
    A.elem( arma::find(A > 0.5) ) += 123.0;

    arma::cube C = A.head_slices(2);  // get first two slices

    A.head_slices(2) += 123.0;

    A.save("A.mat");    // default save format is binary
    

    return(wrap( C ));
'

subcubeExample = cxxfunction(body=subcube, 
                       plugin="RcppArmadillo")


subcubeExample()
```


## iterators

```{r}
require(inline)
require(RcppArmadillo)
 
code <- '
    using namespace Rcpp;
    
    arma::cube X = arma::cube(2,3,4);
    X.ones();


    arma::cube::iterator a = X.begin();
    arma::cube::iterator b = X.end();

    for(arma::cube::iterator i=a; i!=b; ++i)
    {
       arma::cout << *i << arma::endl;
    }

    arma::cube::slice_iterator c = X.begin_slice(1);  // start of slice 1
    arma::cube::slice_iterator d = X.end_slice(2);    // end of slice 2
    
    for(arma::cube::slice_iterator i=c; i!=d; ++i)
    {
        arma::cout << *i << arma::endl;
         (*i) = 123.0;
    }
    

    return(wrap( X ));
'

iterExample = cxxfunction(body=code, 
                       plugin="RcppArmadillo")


iterExample()
```



        #include <RcppArmadillo.h>
        // [[Rcpp::depends(RcppArmadillo)]]
        
        // [[Rcpp::export]]
        arma::cube array2cube( SEXP myArray ) {
        
        Rcpp::NumericVector vecArray(myArray);
        Rcpp::IntegerVector arrayDims = vecArray.attr("dim");
        
        arma::cube cubeArray(vecArray.begin(), arrayDims[0], arrayDims[1], arrayDims[2], false);
        
        return(cubeArray);  
}


src <- '
    using namespace Rcpp;
     
    NumericVector vecArray(myArray);
        IntegerVector arrayDims = vecArray.attr("dim");
        arma::cube cubeArray(vecArray.begin(), 
                             arrayDims[0], arrayDims[1], arrayDims[2], false);
 
        //change one element in the array/cube
        cubeArray(0,0,0) = 518;  
 
    return(wrap(cubeArray));  
'

















