---
title: "R Notebook"
output: html_notebook
---

## Defining Sub-Functions
While writing a bunch of nested loops is great and all, we may also want to be able to write C++ function that can be used by other C++ functions just like we would with R code. To do this, you will need to define your own namespace before you can define sub-functions to be used by your other functions. This `namespace` will give a prefix to any functions you define so C++ can know what you are talking about. In the example below, the `cdf()` function would get called in my Rcpp program by `mjd::cdf()`. These are three functions that calculate the erf, pdf, and cdf of draws from a normal distribution.

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
//[[Rcpp::depends(RcppArmadillo)]]
using namespace Rcpp;

namespace mjd {

    // Returns the erf() of a value (not super precise, but ok)
    double erf(double x)
    {  
     double y = 1.0 / ( 1.0 + 0.3275911 * x);   
     return 1 - (((((
            + 1.061405429  * y
            - 1.453152027) * y
            + 1.421413741) * y
            - 0.284496736) * y 
            + 0.254829592) * y) 
            * exp (-x * x);      
    }

    // Returns the probability of x, given the distribution described by mu and sigma.
    double pdf(double x, double mu, double sigma)
    {
      //Constants
      static const double pi = 3.14159265; 
      return exp( -1 * (x - mu) * (x - mu) / (2 * sigma * sigma)) / (sigma * sqrt(2 * pi));
    }

    // Returns the probability of [-inf,x] of a gaussian distribution
    double cdf(double x, double mu, double sigma)
    {
        return 0.5 * (1 + mjd::erf((x - mu) / (sigma * sqrt(2.))));
    }
}  // end of namespace

// [[Rcpp::export]]
double Mutual_Information(
    double mu,
    double sigma,
    double draw
    ){
        double cdf_val = mjd::cdf(draw, mu, sigma);
        return cdf_val;    
    }
```


We can then call these functions within another C++ function as in the following block of code. Again, note that to actually make the function available to R, we need to include an `// [[Rcpp::export]]` statement on the line directly before the function definition starts.

```{r}
Mutual_Information(3, 0.5, 10)
```

