---
title: "R Notebook"
output: html_notebook
---


http://gallery.rcpp.org/articles/passing-cpp-function-pointers/

To set the stage, let us consider two simple functions of a vector

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
//[[Rcpp::depends(RcppArmadillo)]]


using namespace arma; 
using namespace Rcpp;

vec fun1_cpp(const vec& x) {	// a first function 
    vec y = x + x;
    return (y);
}

vec fun2_cpp(const vec& x) {	// and a second function
    vec y = 10 * x;
    return (y);
}

// This typdef is important and just says that funcPtr will take a const 
// reference to a vec and return a vector – just like our two functions above

    typedef vec (*funcPtr)(const vec& x);

/* 
 * All the functions that have // [[Rcpp::export]]
 * are visible from R
*/  


/* 
 * The following function takes a string argument, picks a function and 
 * returns it wrapped as an external pointer SEXP. We could return this to
 * R as well.
*/

    // [[Rcpp::export]]
    XPtr<funcPtr> putFunPtrInXPtr(std::string fstr) {
        if (fstr == "fun1")
            return(XPtr<funcPtr>(new funcPtr(&fun1_cpp)));  // refers to fun1_cpp
        else if (fstr == "fun2")
            return(XPtr<funcPtr>(new funcPtr(&fun2_cpp)));  // refers to fun2_cpp
        else
            return XPtr<funcPtr>(R_NilValue); // runtime error as NULL no XPtr
    }


// A simple test of this function follows. First a function using it:

    // [[Rcpp::export]]
    vec callViaString(const vec x, std::string funname) {
        XPtr<funcPtr> xpfun = putFunPtrInXPtr(funname);  // get the real C++ function
        funcPtr fun = *xpfun;                            // declare the function pointer
        vec y = fun(x);
        return (y);
    }

    
// And then a call, showing access to both functions:
// callViaXPtr is the function visible from R

    // [[Rcpp::export]]
    vec callViaXPtr(const vec x, SEXP xpsexp) {
        XPtr<funcPtr> xpfun(xpsexp);
        funcPtr fun = *xpfun;
        vec y = fun(x);
        return (y);
    }
```

These are pretty boring and standard functions, and we could simple switch between them via if/else statements. Where it gets interesting is via the SEXP wrapping offered by XPtr below.

But before we get there, let us do this one step at a time.

And then a call, showing access to both functions:

```{r}
# call function fun1
callViaString(1:3, "fun1")
```


```{r}
# call function fun2
callViaString(1:3, "fun2")
```


But more interestingly, we can also receive a function pointer via the SEXP wrapping:

```{r}
fun <- putFunPtrInXPtr("fun1")
```

And use it in this function which no longer switches:
```{r}
callViaXPtr(1:4, fun)
```



