---
title: "R Notebook"
output: html_notebook
---

```{r engine= 'Rcpp'}

// [[Rcpp::depends(RcppArmadillo)]]

#include <RcppArmadillo.h>
#include <Rcpp.h>  

using namespace Rcpp; 

/* 
                         By value
*/ 

// First passing Armadillo matrices by value
void doStuff1(arma::mat c, arma::mat d){
  double x = c(1,1) + d(1,1);
} 

// [[Rcpp::export]]
void twoMatrices1(NumericMatrix A, NumericMatrix B) {
    arma::mat M1(A.begin(), A.nrow(), A.ncol(), false);
    arma::mat M2(B.begin(), B.nrow(), B.ncol(), false);
  
    for (int i = 0; i < 100; i++){
      doStuff1(M1, M2);
    }
} 

/* 
                        By reference
*/ 

// Now passign Armadillo matrices by reference
void doStuff2(const arma::mat& c, const arma::mat& d){
  double x = c(1,1) + d(1,1);
}

// [[Rcpp::export]]
void twoMatrices2(NumericMatrix A, NumericMatrix B) {
    arma::mat M1(A.begin(), A.nrow(), A.ncol(), false);
    arma::mat M2(B.begin(), B.nrow(), B.ncol(), false);
  
    for (int i = 0; i < 100; i++){
      doStuff2(M1, M2);
    }
}

```

Now I am going to call twoMatrices1() and twoMatrices2() from R, using rbenchmark package to compare their speed. I am going to use fairly large matrices (100x100), so the cost of moving them around is significant.

```{r}
library(RcppArmadillo)
library(rbenchmark)
 
# sourceCpp("matrix_code.cpp")
 
x = matrix(1,500,500)
y = matrix(1,500,500)
 
res <- benchmark(twoMatrices1(x, y), twoMatrices2(x, y), 
                 columns = c("test", "replications", "elapsed", "relative"), 
                 order="relative", replications=1000)
print(res)
```
As you can see, passing two 100x100 Armadillo matrices by value is more than 100 times slower than passing these matrices by reference. Perhaps not a very surprising outcome, but an important one to keep in mind when porting R code to C++.
