---
title: "R Notebook"
output: html_notebook
---

Source: http://mcglinn.web.unc.edu/blog/linking-c-with-r-in-windows/

        /*
        Filename: "sequence_examples.c"
        Return a vectors of sequentially summed values
        Arguments:
        start -- value to start the sum at
        size -- the number of elements to return
        sumVect -- the vector of summed output values
        */
        
        void sumSeq(int *start, int *size, int *sumVect){
            /*
            This function provides a simple sequential sum
            where F[n] = F[n-1] + n
            */
            int i, j ;
            j = 0 ;
            for(i = *start; i < (*start + *size); i++){
                if(i == *start){
                    sumVect[j] = i ;
                }
                else{
                    sumVect[j] = sumVect[j-1] + i ;
                }
                j ++ ;
            }
        }
        
        void fiboSeq(int *size, int *sumVect){
            /*
            This function returns the Fibonacci sequence
            where F[n] = F[n-1] + F[n-2]
            */
            int i ;
            sumVect[0] = 0 ;
            sumVect[1] = 1 ;
            for(i = 2; i < *size; i++){
                sumVect[i] = sumVect[i-1] + sumVect[i-2] ;
            }
        }


```{r}
# compile the C code
system("R CMD SHLIB ../src/sequence_examples.c")
```

```{r}
# this loads the library in any system, Windows or Linux
dyn.load(paste("../src/sequence_examples", .Platform$dynlib.ext, sep = ""))
```


```{r}
.C("sumSeq", start = as.integer(10), size = as.integer(5),
   sumVect = as.integer(rep(0, 5)))

.C("fiboSeq", size = as.integer(5),
   sumVect = as.integer(rep(0, 5)))
```

These R commands should return lists that contain the input to the C function and the resulting output. Because we are only usually interested in the output it is a good idea to be more specific about what portion of the list you would like returned.


```{r}
.C("sumSeq", start = as.integer(10), size = as.integer(5),
   sumVect = as.integer(rep(0, 5)))$sumVect

.C("fiboSeq", size = as.integer(5),
   sumVect = as.integer(rep(0, 5)))$sumVect
```

