---
title: "R Notebook"
output: html_notebook
---

```{r}
# fooblas.c will not compile from within RStudio
# it has to be compiled from the command line
# $    R CMD SHLIB fooblas.c
system("R CMD SHLIB ../src/fooblas.c")
```


```{r}
# this loads the library in any system, Windows or Linux
dyn.load(paste("../src/fooblas", .Platform$dynlib.ext, sep = ""))
```


```{r}
# void foo(int *dimension, double *vect1, double *vect2)

# wrapper
c.fooblas = function(a, b) {
   res <- .C('fooblas',
      as.integer(length(a)),
      as.double(a),
      as.double(b))
   res[c(2,3)]
}
```


```{r}
c.fooblas(c(5, 10), c(3,6))
```

```{r}
c.fooblas(matrix(5, nrow=3, ncol=2), matrix(2, nrow=3, ncol=2))
```

## another version of swap using matrices

```{r}
# void foo(int *dimension, double *vect1, double *vect2)
# dimension: number of elements in input vector(s)

c.fooblas2 = function(a, b) {
    sa <- sb <- a
   res <- .C('fooblas',
      as.integer(length(a)),
      as.double(a),
      as.double(b))
   sa[] <- res[[2]]
   sb[] <- res[[3]]
   list(a, b, sa, sb)  # return original and swapped
}
```


```{r}
c.fooblas2(matrix(5, nrow=3, ncol=2), matrix(2, nrow=3, ncol=2))
```

```{r}
c.fooblas2(array(4, dim=c(4,3,2)), array(7, dim = c(4,3,2)))
```


```{r}
# this loads the library in any system, Windows or Linux
dyn.unload(paste("../src/fooblas", .Platform$dynlib.ext, sep = ""))
```

## Notes about compilation

* You need a file Makevars in ./src
* fooblas.c will not compile from within RStudio. It has to be compiled from the command line like this: `R CMD SHLIB fooblas.c`
*

> after removing Makevars in ./src

    msfz751@HPZBOOK15G2W10 MSYS /i/pub.packages/rBlaster/src
    $ R CMD SHLIB fooblas.c
    c:/Rtools/mingw_64/bin/gcc  -I"C:/PROGRA~1/R/R-34~1.1/include" -DNDEBUG     -I"d:/Compiler/gcc-4.9.3/local330/include"     -O2 -Wall  -std=gnu99 -mtune=core2 -c fooblas.c -o fooblas.o
    c:/Rtools/mingw_64/bin/gcc -shared -s -static-libgcc -o fooblas.dll tmp.def fooblas.o -Ld:/Compiler/gcc-4.9.3/local330/lib/x64 -Ld:/Compiler/gcc-4.9.3/local330/lib -LC:/PROGRA~1/R/R-34~1.1/bin/x64 -lR
    fooblas.o:fooblas.c:(.text+0x28): undefined reference to `dswap_'
    collect2.exe: error: ld returned 1 exit status

    
> after replacing Makevars back to ./src

    msfz751@HPZBOOK15G2W10 MSYS /i/pub.packages/rBlaster/src
    $ R CMD SHLIB fooblas.c
    c:/Rtools/mingw_64/bin/gcc -shared -s -static-libgcc -o fooblas.dll tmp.def fooblas.o -fopenmp -LC:/PROGRA~1/R/R-34~1.1/bin/x64 -lRlapack -LC:/PROGRA~1/R/R-34~1.1/bin/x64 -lRblas -lgfortran -lm -lquadmath -Ld:/Compiler/gcc-4.9.3/local330/lib/x64 -Ld:/Compiler/gcc-4.9.3/local330/lib -LC:/PROGRA~1/R/R-34~1.1/bin/x64 -lR



