---
title: "Kalman test"
output: html_notebook
---

https://github.com/RcppCore/RcppArmadillo/tree/master/inst/examples/kalman

PDF: http://dirk.eddelbuettel.com/papers/RcppArmadillo-intro.pdf

```{r}
# direct translation of the original Matlab implementation

FirstKalmanR <- function(pos) {

    kalmanfilter <- function(z) {
        dt <- 1
        A <- matrix(c( 1, 0, dt, 0, 0, 0, 0, 1, 0, dt, 0, 0,  # x,  y
                       0, 0, 1, 0, dt, 0, 0, 0, 0, 1, 0, dt,  # Vx, Vy
                       0, 0, 0, 0, 1,  0, 0, 0, 0, 0, 0,  1), # Ax, Ay
                    6, 6, byrow=TRUE)
        H <- matrix( c(1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0),
                    2, 6, byrow=TRUE)
        Q <- diag(6)
        R <- 1000 * diag(2)

        xprd <- A %*% xest		    # predicted state and covriance
        pprd <- A %*% pest %*% t(A) + Q

        S <- H %*% t(pprd) %*% t(H) + R     # estimation
        B <- H %*% t(pprd)
        kalmangain <- t(solve(S, B))

        ## estimated state and covariance, assign to vars in parent env
        xest <<- xprd + kalmangain %*% (z - H %*% xprd)
        pest <<- pprd - kalmangain %*% H %*% pprd

        ## compute the estimated measurements
        y <- H %*% xest
    }
    xest <- matrix(0, 6, 1)
    pest <- matrix(0, 6, 6)

    N <- nrow(pos)
    y <- matrix(NA, N, 2)
    for (i in 1:N) {
        y[i,] <- kalmanfilter(t(pos[i,,drop=FALSE]))
    }
    invisible(y)
}



```


```{r}
# R implementaton of Kalman filter
KalmanR <- function(pos) {

    kalmanfilter <- function(z) {
        ## predicted state and covariance
        xprd <- A %*% xest
        pprd <- A %*% pest %*% t(A) + Q

        ## estimation
        S <- H %*% t(pprd) %*% t(H) + R
        B <- H %*% t(pprd)

        kalmangain <- t(solve(S, B))

        ## estimated state and covariance
        ## assigned to vars in parent env
        xest <<- xprd + kalmangain %*% (z - H %*% xprd)
        pest <<- pprd - kalmangain %*% H %*% pprd

        ## compute the estimated measurements
        y <- H %*% xest
    }

    dt <- 1
    A <- matrix( c( 1, 0, dt, 0, 0, 0,  # x
                   0, 1, 0, dt, 0, 0,   # y
                   0, 0, 1, 0, dt, 0,   # Vx
                   0, 0, 0, 1, 0, dt,   # Vy
                   0, 0, 0, 0, 1,  0,   # Ax
                   0, 0, 0, 0, 0,  1),  # Ay
                6, 6, byrow=TRUE)
    H <- matrix( c(1, 0, 0, 0, 0, 0,
                   0, 1, 0, 0, 0, 0),
                2, 6, byrow=TRUE)
    Q <- diag(6)
    R <- 1000 * diag(2)
    N <- nrow(pos)
    Y <- matrix(NA, N, 2)

    xest <- matrix(0, 6, 1)
    pest <- matrix(0, 6, 6)

    for (i in 1:N) {
        Y[i,] <- kalmanfilter(t(pos[i,,drop=FALSE]))
    }
    invisible(Y)
}
```


```{r}
## Improved version supplied by one referee, not shown in paper
KalmanRimp <- function(pos) {

    kalmanfilter <- function(z) {
        ## predicted state and covriance
        xprd <- xest %*% A
        pprd <- crossprod(pest %*% A, A) + Q

        ## estimation
        B <- crossprod(H, pprd)
        S <- B %*% H + R

        ##  kalmangain <- (S \ B)'
        kalmangain <- solve(S, B)

        ## estimated state and covariance, assign to vars in parent env
        xest <<- xprd + (z - xprd %*% H) %*% kalmangain
        pest <<- pprd - pprd %*% H %*% kalmangain

        ## compute the estimated measurements
        y <- xest %*% H
    }

    dt <- 1
    A <- matrix( c( 1, 0, dt, 0, 0, 0,  # x
                   0, 1, 0, dt, 0, 0,   # y
                   0, 0, 1, 0, dt, 0,   # Vx
                   0, 0, 0, 1, 0, dt,   # Vy
                   0, 0, 0, 0, 1,  0,   # Ax
                   0, 0, 0, 0, 0,  1),  # Ay
                6, 6, byrow=FALSE)
    
    H <- matrix( c(1, 0, 0, 0, 0, 0,
                   0, 1, 0, 0, 0, 0),
                6, 2, byrow=FALSE)
    Q <- diag(6)
    R <- 1000 * diag(2)

    N <- nrow(pos)
    y <- matrix(NA, N, 2)

    xest <- matrix(0, 1, 6)
    pest <- matrix(0, 6, 6)

    for (i in 1:N) {
        y[i,] <- kalmanfilter(pos[i,,drop=FALSE])
    }

    invisible(y)
}
```

### Listing 3

```{r}
# kalman Cpp implemented in Armadillo

kalmanClass <- '

// [[Rcpp::depends(RcppArmadillo)]]

#include <RcppArmadillo.h>

using namespace arma;

class Kalman {
private:
    mat A, H, Q, R, xest, pest;
    double dt;

public:
    // constructor, sets up data structures
    Kalman() : dt(1.0) {
        A.eye(6,6);
        A(0,2) = A(1,3) = A(2,4) = A(3,5) = dt;
        H.zeros(2,6);
        H(0,0) = H(1,1) = 1.0;
        Q.eye(6,6);
        R = 1000 * eye(2,2);
        xest.zeros(6,1);
        pest.zeros(6,6);
    }

    // sole member function: estimate model
    mat estimate(const mat & Z) {
       unsigned int n = Z.n_rows, k = Z.n_cols;
       mat Y = zeros(n, k);
       mat xprd, pprd, S, B, kalmangain;
       colvec z, y;

       for (unsigned int i = 0; i<n; i++) {
           z = Z.row(i).t();
           // predicted state and covariance
           xprd = A * xest;
           pprd = A * pest * A.t() + Q;
           // estimation
           S = H * pprd.t() * H.t() + R;
           B = H * pprd.t();
           kalmangain = (solve(S, B)).t();
           // estimated state and covariance
           xest = xprd + kalmangain * (z - H * xprd);
           pest = pprd - kalmangain * H * pprd;
           // compute the estimated measurements
           y = H * xest;
           Y.row(i) = y.t();
       }
       return Y;
    }
};


// [[Rcpp::export]]
mat KalmanCpp(mat Z) {
  Kalman K;
  mat Y = K.estimate(Z);
  return Y;
}

'
```


### Listing 4
```{r}
library(RcppArmadillo)
library(inline)

kalmanSrc <- '
 mat Z = as<mat>(ZS );   // p assed from R
 Kalman K;
 mat Y = K.estimate(Z) ;
 return wrap(Y) ;
'

KalmanCpp <- cxxfunction(signature(ZS="numeric"), 
                         body=kalmanSrc ,include=kalmanClass, 
                         plugin="RcppArmadillo")
```





```{r}
require( rbenchmark )
require( compiler )

# FirstKalmanR.R: direct translation of the original Matlab implementation

# source("FirstKalmanR.R")
# source("KalmanR.R")
# source("KalmanRimp.R")
# Rcpp::sourceCpp("Kalman.cpp")

FirstKalmanRC <- cmpfun(FirstKalmanR) # byte-compiled version of FirstKalmanRC
KalmanRC <- cmpfun(KalmanR)           # byte-compiled version of KalmanR
KalmanRimpC <- cmpfun(KalmanRimp) ## Improved version by referee, not shown in paper


## Read data, ensure identical results
pos <- as.matrix(read.table("pos.txt", header=FALSE,
                            col.names=c("x","y")))

stopifnot(all.equal(KalmanR(pos), KalmanRC(pos)),
          all.equal(KalmanRC(pos), KalmanCpp(pos)),
          all.equal(KalmanCpp(pos), FirstKalmanRC(pos)),
          all.equal(KalmanCpp(pos), FirstKalmanR(pos)),
          all.equal(KalmanCpp(pos), KalmanRimp(pos)),
          all.equal(KalmanCpp(pos), KalmanRimpC(pos)))

# KalmanCpp is the RcppArmadillo based implementation in C++

res <- benchmark(KalmanR(pos), KalmanRC(pos),
                 KalmanRimp(pos), KalmanRimpC(pos),
                 FirstKalmanR(pos), FirstKalmanRC(pos),
                 KalmanCpp(pos),
                 columns = c("test", "replications",
                             "elapsed", "relative"),
                 order="relative",
                 replications=500)

print(res[,1:4])

```

KalmanCpp is the RcppArmadillo based implementation in C++ shown in Listings 3 and 4. KalmanR is the R implementation shown in Listing 2; KalmanRC is the byte-compiled version of KalmanR. FirstKalmanR is a direct translation of the original Matlab implementation mentioned in Section 4; FirstKalmanRC is the corresponding byte-compiled version. Timings are averaged over 100 replications. The comparison was made using R version 2.15.1, Rcpp version 0.9.12 and RcppArmadillo version 0.3.2.0 on Ubuntu 12.04 running in 64-bit mode on a 2.67 GHz Intel i7 processor.

A small example implementing a Kalman filter illustrated two key aspects. First,
orders of magnitude of performance gains can be obtained by deploying C++ code
along with R. Second, the ease of use and readability of the corresponding C++
code is similar to the R code from which it was derived. This combination makes
RcppArmadillo a compelling tool in the arsenal of applied researchers deploying
computational methods in statistical computing and data analysis.

