---
title: "R Notebook"
output: html_notebook
---

http://lists.r-forge.r-project.org/pipermail/rcpp-devel/2016-September/009375.html

## Looping a cube

It is much, much, much easier to deal with small, simplified,
examples, as Dirk and others have explained before. Changing your code
to make n = 4 and the values seq(1, 12), we can see the pattern. For
each slice, you are taking the product down the rows of "slice" number
of entries. So the first slice is the 1-element product of the
entries, or just a copy of the original matrix. 

Now, putting the index
names back in familiar context (M rows and N columns), for the second
slice, C(2, 1, 2) = A(1, 1) * A(2, 1). C(3, 1, 2) = A(2, 1) * A(3, 1),
etc. For the third slice, we have 3-element products So C(1, , 3) and
C(2, , 3) are 0. C(3, 1, 3) = A(1, 1) * A(2, 1) * A(3, 1), etc. In the
last slice there is only one entry in the last row and it is the
product of each column in its entirety.

That being said, the way I would approach it is to use Armadillo's
subsetting to get the needed vector and use Armadillo's product
function to get their product.

    #include <RcppArmadillo.h>
    using namespace Rcpp;
    using namespace RcppArmadillo;
    using namespace arma;
    //[[Rcpp::depends(RcppArmadillo)]]
    
    //[[Rcpp::export]]
    cube up_C(mat a){
        int m = a.n_rows;
        int n = a.n_cols;
        
        cube C = cube(m, n, m, fill::zeros);
        
        C.slice(0) = a;
        for (int slices = 1; slices < m; ++slices){
            for (int j = 0; j < n; ++j) {
                for (int i = slices; i < m; ++i) {
                    vec S = C.subcube(i - slices, j, 0, i, j, 0);
                    C(i, j, slices) = prod(S);
                }
            }
        }
        return(C);
    }

```{r}
library(Rcpp)

sourceCpp("../src/armaLoop.cpp")

mx <- matrix(10, nrow=5, ncol=2)    # R array to send
up_C(mx)
```

This pulls the "slices" length subset of the jth column starting at
slices and ending at i from the 0th slice (which is the original
matrix) and puts the product in to the C(i, j, slices) entry.

You can test the two are identical through identical(up_C(a), up_R(a))
for any a, and the C version is, of course, much faster especially for
large matrices.

It often pays to start with something simpler, and see if there is a
pattern you can identify which would help you conceptually, even if
not completely optimized.


```{r}
# the R equivalent above

up_R <- function(a){
  m = dim(a)[[1]]
  n = dim(a)[[2]]
  
  b = array(as.double(0), dim = c(m, n, m))
  
  for(i in 1:m){
    for(j in 1:n){
      for(q in 1:i){
        if(q==1){
          b[i, j, q] = a[i, j];
        }
        else{
          b[i, j, q] = b[i-1, j, q-1] * a[i, j];
        }
      }
    }
  }
  return(b)
}

up_R(mx)
```



```{r}
library(Rcpp)

src <- '
#include <RcppArmadillo.h>
using namespace Rcpp;
using namespace RcppArmadillo;
using namespace arma;
//[[Rcpp::depends(RcppArmadillo)]]

//[[Rcpp::export]]
cube up_CI(mat a){
    int m = a.n_rows;    // get rows from matrix a
    int n = a.n_cols;    // get cols from matrix a
    
    cube C = cube(m, n, m, fill::zeros);  // create array and fill w/zeros
    
    C.slice(0) = a;  // output array 1st matrix is matrix a
    
    for (int slices = 1; slices < m; ++slices){   // move through slices
        for (int j = 0; j < n; ++j) {             // move through cols
            for (int i = slices; i < m; ++i) {    // move through rows
                // extract subcubes
                vec S = C.subcube(i - slices, j, 0, i, j, 0);
                // Armadillo product function arma::prod(S)
                C(i, j, slices) = prod(S);
            }
        }
    }
    return(C);
}
'

cppFunction(code=src, depends="RcppArmadillo")

mx <- matrix(10, nrow=5, ncol=2)    # R array to send
up_CI(mx)

```









