---
title: "R Notebook"
output: html_notebook
---

## Displace a row down in array matrices

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp;
using namespace RcppArmadillo;
using namespace arma;

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

cube up_CI(mat a){
    int m = a.n_rows;    // get rows from matrix a
    int n = a.n_cols;    // get cols from matrix a
    
    cube C = cube(m, n, m, fill::zeros);  // create array and fill w/zeros
    
    C.slice(0) = a;  // output array 1st matrix is matrix a
    
    for (int slices = 1; slices < m; ++slices){   // move through slices
        for (int j = 0; j < n; ++j) {             // move through cols
            for (int i = slices; i < m; ++i) {    // move through rows
                // extract subcubes
                vec S = C.subcube(i - slices, j, 0, i, j, 0);
                // Armadillo product function arma::prod(S)
                C(i, j, slices) = prod(S);
            }
        }
    }
    return(C);
}


```

```{r}
mx <- matrix(10, nrow=5, ncol=2)    # R array to send
up_CI(mx)
```


## Multivariate normal

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

using namespace Rcpp;

// [[Rcpp::export]]
arma::mat mvrnormArma(int n, arma::vec mu, arma::mat sigma) {
   int ncols = sigma.n_cols;
   arma::mat Y = arma::randn(n, ncols);
   return arma::repmat(mu, 1, n).t() + Y * arma::chol(sigma);
}
```


```{r}
sigma <- matrix(c(1, 0.9, -0.3, 0.9, 1, -0.4, -0.3, -0.4, 1), ncol = 3)
mu <- c(10, 5, -3)

mvrnormArma(50, mu, sigma)

colMeans(mvrnormArma(50, mu, sigma))
```


## Join slices of two cubes/arrays

```{r engine='Rcpp'}
/* for two cubes C and D: join the slices of C with the slices of D; cubes C 
 * and D must have the same number of rows and columns (ie. all slices must 
 * have the same size)
 */ 
    
#include <RcppArmadillo.h>
using namespace Rcpp;
using namespace RcppArmadillo;
using namespace arma;

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

cube joinSlices( cube c1, cube c2 ) {


    //int ns = c1.n_slices;
    cout << "cube c1 has " << c1.n_slices << " slices" << endl;
    cout << "cube c2 has " << c2.n_slices << " slices" << endl;

    cube out = join_slices(c1, c2);
    
    cout << "cube out has " << out.n_slices << " slices" << endl;
    return out;
}
```


```{r}
a1 <- array(1:24, dim = c(4,3,2))
a2 <- array(25:48, dim = c(4,3,2))

joinSlices(a1, a2)
```

## column means of a cube (colMeans)

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace arma;

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

mat cubeColMeans(arma::cube X){
    int nSlice = X.n_slices;
    int nCol = X.n_cols;
    int nRow = X.n_rows;
    
    arma::vec Vtmp(nCol);
    arma::mat Mtmp(nRow, nCol);
    arma::mat Means(nCol, nSlice);
    
    for (int i = 0; i < nSlice; i++){
        Mtmp = X.slice(i);
        for(int j = 0; j < nCol; j++){
            Vtmp(j) = sum(Mtmp.col(j)) / nRow; 
            // Rcout << "Vtmp: \n"   << Vtmp(j) << std::endl ;
        }
        
        Means.col(i) = Vtmp;
        // Rcout << "Means.col: \n"   << Means.col(i) << std::endl ;
        
    }
    return(Means);
}

```


```{r}
xl <- array(1:100, c(5, 10, 2))
# xl


# R equivalent function
# rcubeColMeans <- function(x) t(apply(x, 2, colMeans)) # another way
rcubeColMeans <- function(x) apply(x, c(2,3), mean)

identical(rcubeColMeans(xl), cubeColMeans(xl))

```

## rowSums

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace arma;

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

mat cubeRowMeans(arma::cube X){
    int nSlice = X.n_slices;
    int nCol = X.n_cols;
    int nRow = X.n_rows;
    
    arma::vec Vtmp(nRow);          // vector with size of the rows
    arma::mat Mtmp(nCol, nRow);    // extraction matrix with as-is dimensions
    arma::mat Means(nRow, nSlice); // matrix with size nRow by number of slices
    
    for (int k = 0; k < nSlice; k++){
        Mtmp = X.slice(k);
        // Rcout << "Matrix: \n"   << Mtmp << std::endl ;
        for(int i = 0; i < nRow; i++){
            Vtmp(i) = sum(Mtmp.row(i)) / nCol; 
            // Rcout << "Vtmp: \n"   << Vtmp(i) << std::endl ;
        }
        Means.col(k) = Vtmp;
        //Rcout << "Means.col: \n"   << Means.col(k) << std::endl ;
    }
    return(Means);
}

```


```{r}
xl <- array(1:100, c(5, 10, 2))
# xl

# R equivalent function
rcubeRowMeans <- function(x) apply(x, c(1,3), mean)
identical(rcubeRowMeans(xl), cubeRowMeans(xl))
```


## add a matrix slice to a cube

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace RcppArmadillo;
using namespace arma;


//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

cube add_matrix_slice(cube c1, mat m1) {

    // Normal Matrix dimensions
    int xdim = c1.n_rows;
    int ydim = c1.n_cols;
    int tdim = c1.n_slices;  // Depth of Array


    cout << "cube c1 had before " << c1.n_slices << " slices" << endl;

    // resulting cube
    cube out(c1.begin(), xdim, ydim, tdim+1);

    // reshape matrix to size of cube slices
    mat rm1 = reshape(m1, xdim, ydim);         // this is critical !!!

    // join slice of cube plus reshaped matrix
    out = join_slices(c1, rm1);
    
    cout << "cube out has " << out.n_slices << " slices now" << endl;
    return out;
}    
```


```{r}
m1 <- matrix(100:115,  nr=4)           # matrix is 4x4
a1 <- array(1:24, dim = c(4,3,2))

res  <- add_matrix_slice(a1, m1)
res

test <- array(c(a1, m1), dim = c(4,3,3))
all.equal(test, res)
```


## add a shifted matrix slice to a cube

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace RcppArmadillo;
using namespace arma;


//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

cube add_matrix_shift_slice(cube c1, mat m1, int N=0, int dim=0) {

    // Normal Matrix dimensions
    int xdim = c1.n_rows;
    int ydim = c1.n_cols;
    int tdim = c1.n_slices;  // Depth of Array


    cout << "cube c1 had before " << c1.n_slices << " slices" << endl;

    // resulting cube
    cube out(c1.begin(), xdim, ydim, tdim+1);
     
    /* arma::shift(X, N, dim)
     * For matrix X, generate a copy of the matrix with the elements shifted by 
     * N positions in each column (dim=0), or each row (dim=1)
     * N can be positive or negative
     * The dim argument is optional; by default dim=0 is used     
     */
    mat rm1 = shift(m1, N, dim);
    // Rcout << "rm1: \n"   << rm1 << std::endl ;

    // reshape matrix to size of cube slices
    rm1 = reshape(rm1, xdim, ydim);         // this is critical !!!

    // join slice of cube plus reshaped matrix
    out = join_slices(c1, rm1);
    
    cout << "cube out has " << out.n_slices << " slices now" << endl;
    return out;
}    
```


```{r}
m1 <- matrix(100:199,  nr=10)           # matrix is 4x4
m1
a1 <- array(1:24, dim = c(4,3,2))

cat("no shift, default values \n")
add_matrix_shift_slice(a1, m1)

# shift 2 columns to the left
cat("shift 2 columns to the left \n")
add_matrix_shift_slice(a1, m1, N = 2, dim = 1)

# shift 6 columns to the right
cat("shift 6 columns to the right \n")
add_matrix_shift_slice(a1, m1, N = -6, dim = 1)
```



```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace RcppArmadillo;
using namespace arma;


//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

cube extract_cube(cube c1, vec v1, vec v2) {
    
    cube out = c1;
    //c1.print("c1:");
    
    //v1.each( [](vec& a){ a.print(); } );     // lambda function with non-const vector

    cout << "vector v1 has " << v1 << "values" << endl;
    cout << "vector v1 has " << v1.size() << " elements" << endl;
    cout << "vector v2 has " << v2.size() << " elements" << endl;


    // join slice of cube plus reshaped matrix
    out = c1.subcube(v1[0],v1[1],v1[2],  v2[0], v2[1], v2[2]);
    
    //cout << "cube out has " << out.n_slices << " slices" << endl;
    return out;
}
```


```{r}
a1 <- array(1:24, dim = c(4,3,2))
a1
u1 <- c(0,0,0)   # from
u2 <- c(1,1,1)   # to
res  <- extract_cube(a1, u1, u2)
res
```

## calculates the mutual information of an arbitrary joint distribution.
http://www.mjdenny.com/Rcpp_Intro.html

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

double Mutual_Information(arma::mat joint_dist) {
    
    joint_dist = joint_dist/sum(sum(joint_dist));
    double mutual_information = 0;
    int num_rows = joint_dist.n_rows;
    int num_cols = joint_dist.n_cols;
    arma::mat colsums = sum(joint_dist,0);
    arma::mat rowsums = sum(joint_dist,1);
    for(int i = 0; i < num_rows; ++i){
        for(int j = 0; j <  num_cols; ++j){
            double temp = log((joint_dist(i,j)/(colsums[j]*rowsums[i])));
            if(!std::isfinite(temp)){
                temp = 0;
            }
            mutual_information += joint_dist(i,j) * temp; 
        }
    } 
    return mutual_information;    
}
```


```{r}
Mutual_Information(matrix(1:600, nrow = 10))
```


## rdirichlet in Rcpp -- 3/16/16
http://www.mjdenny.com/blog.html#4-5-15

A paper (and R package) I am currently working on requires me to sample quite a bit from a Dirichlet distribution. There are a number of nice options for doing this in many languages, including R, but as far as I am aware, none of these interface nicely with Rcpp (and are easily portable in an R package). While browsing StackExchange and Dirk Eddelbuettel's website, hoping to find a magical solution, I came across a link to the Wikipedia page on the Dirichlet_distribution. It points to a pretty simple way to sample from a Dirichlet distribution using a bunch of Gammas, which Rcpp does provide, so I decided to code it up myself! Lets take a look at the source code :


```{r engine='Rcpp'}
// [[Rcpp::depends(RcppArmadillo)]]
#include <RcppArmadillo.h>

// [[Rcpp::export]]
arma::mat rdirichlet_cpp(int num_samples,
                         arma::vec alpha_m) {
    int distribution_size = alpha_m.n_elem;
    // each row will be a draw from a Dirichlet
    arma::mat distribution = arma::zeros(num_samples, distribution_size);

    for (int i = 0; i < num_samples; ++i) {
        double sum_term = 0;
        // loop through the distribution and draw Gamma variables
        for (int j = 0; j < distribution_size; ++j) {
            double cur = R::rgamma(alpha_m[j],1.0);
            distribution(i,j) = cur;
            sum_term += cur;
        }
        // now normalize
        for (int j = 0; j < distribution_size; ++j) {
            distribution(i,j) = distribution(i,j)/sum_term;
        }
    }
    return(distribution);
}
```


I decided to try this out against the rdirichlet() functions included in the gtools and MCMCpack packages. It looks like about a 25% speedup. The more important thing is that this code is easy to stick in your Rcpp project, which makes for much faster sampling.

```{r}
alpha <- 5
topics <- 100
alpha_m <- rep(alpha/topics,topics)
 
system.time({
     temp <- gtools::rdirichlet(1000000, alpha_m)
})

 
system.time({
     temp2 <- MCMCpack::rdirichlet(1000000, alpha_m)
 })

system.time({
     temp2 <- rdirichlet_cpp(1000000, alpha_m)
 })


```

Which produces a plot that should look something like the following:

```{r}
alpha <- 5
draws <- 1
dimen <- 20
alpha_m <- rep(alpha/dimen, dimen)

x <- rdirichlet_cpp(draws,alpha_m)

 dat <- data.frame(Category = factor(1:dimen),
                   Density = as.vector(t(x)))

 library(ggplot2)
 ggplot(dat,aes(x = Category,y = Density,ymin = 0, ymax = Density)) +
     geom_point(colour = "darkblue",fill = "darkblue") +
     geom_linerange(colour = "darkblue") +
     scale_y_continuous(lim=c(0,0.5))
```


## QR decomposition
> This is not giving the same results. Programmatically works.

https://stackoverflow.com/questions/10964047/qr-decomposition-in-rcpparmadillo

```{r engine='Rcpp'}
#include <RcppArmadillo.h>
using namespace Rcpp; 
using namespace RcppArmadillo;
using namespace arma;

//[[Rcpp::depends(RcppArmadillo)]]
//[[Rcpp::export]]

Rcpp::List QR(mat Xr, int q ) {
    arma::arma_rng::set_seed(42);
    
    int n = Xr.n_rows, k = Xr.n_cols;
    arma::mat X(Xr.begin(), n, k, false);
    //cout << "X" << X.n_rows << X.n_cols  << endl;

    arma::mat G, Y, B;

    G = arma::randn(q, n);
    //cout << "G \n" << G.n_rows << "x" << G.n_cols  << endl;

    Y = X * G;

    arma::mat Q, R;
    arma::qr(Q, R, Y);

    return Rcpp::List::create(Rcpp::Named("Q")=Q, 
                              Rcpp::Named("R")=R,
                              Rcpp::Named("Y")=Y);
    //return Rcpp::List::create(Rcpp::Named("Y") = Y);
}
```

```{r}
set.seed(42)
mx <- matrix(1:2000, nrow = 1000, ncol = 20)
qr <- QR(mx, 20)
names(qr)
length(qr[[1]])
length(qr[[2]])
length(qr[[3]])

arma.qr <- qr[[1]]
```

```{r}
set.seed(42)
r.qr = qr.Q(qr(mx), complete=TRUE)
identical(r.qr, arma.qr)
```


