---
title: "Rcpp attributes"
output:
  html_document: default
  html_notebook: default
---

A recent addition to Rcpp provides an even more direct connection between C++
and R. This feature is called “attributes” as it is inspired by a C++ extension of
the same name in the new C++11 standard (which will be available to R users only
when CRAN permits use of these extension, which may be years away).

Simply put, “Rcpp attributes” internalizes key features of the inline package
while at the same time reusing some of the infrastructure built for use by inline
such as the plugins.

“Rcpp attributes” adds new functions sourceCpp to source a C++ function
(similar to how sourceis used for R code), cppFunctionfor a similar creation
of a function from a character argument, evalCppfor a direct evaluation of a C++
expression and more.

Behind the scenes, these functions make use of the existing wrappers as<>and
wrap and do in fact rely heavily on them: any arguments with existing convert-
ers to or from SEXP types can be used. The standard build commands such as R
RMD COMPILEand R CMD SHLIB are executed behind the scenes, and template
programming is used to provide compile-time bindings and conversion.
An example may illustrate this:

```{r}
library(Rcpp)
library(inline)

# Listing 2.9
 cpptxt <- '
    int fibonacci(const int x) {
        if (x < 2) return(x);
    return (fibonacci(x - 1)) + fibonacci(x - 2);
    }'

fibCpp <- cppFunction(cpptxt) # compiles, load, links, ..
fibCpp(6)
```

cppFunction returns an R function which calls a wrapper, also created by
cppFunctionin a temporary ﬁle which it also builds. The wrapper function in
turn calls the C++ function we passed as a character string. The build process
administered by cppFunction uses a caching mechanism which ensures that
only one compilation is needed per session (as long as the source code used is un-
changed).

Alternatively, we could pass the name of a ﬁle containing the code to the function
sourceCpp which would compile, link, and load the corresponding C++ code
and assign it to the R function on the left-hand side of the assignment.


```{r}
library(RcppGSL)
 code <- '

     #include <gsl/gsl_vector.h> 
    #include <gsl/gsl_const_mksa.h> // decl of constants
    
    // [[Rcpp::depends(RcppGSL)]]
    std::vector<double> volumes() {
    std::vector<double> v(5);

    v[0] = GSL_CONST_MKSA_US_GALLON; // 1 US gallon
    v[1] = GSL_CONST_MKSA_CANADIAN_GALLON; // 1 Canadian gallon
    v[2] = GSL_CONST_MKSA_UK_GALLON; // 1 UK gallon
    v[3] = GSL_CONST_MKSA_QUART; // 1 quart
    v[4] = GSL_CONST_MKSA_PINT; // 1 pint
    return v;
    }
'

gslVolumes <- cppFunction(code, depends="RcppGSL")
gslVolumes()
```

