---
title: "R Notebook"
output: html_notebook
---

Source: http://adv-r.had.co.nz/C-interface.html

## C data structures
At the C-level, all R objects are stored in a common datatype, the SEXP, or S-expression. All R objects are S-expressions so every C function that you create must return a SEXP as output and take SEXPs as inputs. (Technically, this is a pointer to a structure with typedef SEXPREC.) A SEXP is a variant type, with subtypes for all R’s data structures. The most important types are:

    REALSXP: numeric vector
    INTSXP: integer vector
    LGLSXP: logical vector
    STRSXP: character vector
    VECSXP: list
    CLOSXP: function (closure)
    ENVSXP: environment

```{r}
# compile the C code
system("R CMD SHLIB ../src/add.c")
```

```{r}
# this loads the library in any system, Windows or Linux
dyn.load(paste("../src/add", .Platform$dynlib.ext, sep = ""))
```

```{r}
# In R ----------------------------------------
add <- function(a, b) {
  .Call("add", a, b)
}
```

```{r}
add(3, 7)
```



## inline version

```{r}
library(inline)

add <- cfunction(c(a = "integer", b = "integer"), "
  SEXP result = PROTECT(allocVector(REALSXP, 1));
  REAL(result)[0] = asReal(a) + asReal(b);
  UNPROTECT(1);

  return result;
")

add(1, 5)
```

