---
title: "R Notebook"
output: html_notebook
---


```{r}
# remove the older compiled file
name <- "convolve1D"
cname <- paste0(name, ".cpp")
wkdir <-  "../src"
filename  <- paste(wkdir, name, sep = "/")
cfilename <- paste(wkdir, cname, sep = "/")

file.remove(paste(filename, .Platform$dynlib.ext, sep = ""))
```

```{r}
library(Rcpp)
# compile the C code
system("R CMD SHLIB ../src/convolve1D.cpp")
```


```{r}
# this loads the library in any system, Windows or Linux
dyn.load(paste(filename, .Platform$dynlib.ext, sep = ""))
```



```{r}
# wrapper function to invoke cconv 
rconv <- function(lag, x) {
        .C("cconv", 
           as.integer(lag),
           as.double(x),
           as.integer(length(x)),
           as.double(0.0))[[4]]
}
```

"Observe that we convert all arguments to the approriate type (just to be sure). The .C function returns a list with all its arguments, we only need the last (fourth) argument. In this case we did not name the last argument but extracted it using a numeric index. Whether one uses named arguments (as in Section 5.1) or numeric indices is a matter of personal preference.""

```{r}
x <- rnorm(100)
rconv(0, x)
```

```{r}
x <- rnorm(100)
rconv(1, x)
```

With a little bit more trouble we can vectorize this, 
i.e. make it accept a vector of lags and return a vector of 
convolutions of the same length. The C code now becomes:

    void cconv (int *l, int *m, double *x, int *n, double *s)
    {
      double *y, *z, *u;
      int i;
      for (i = 0; i < *m; i++)
        {
          y = x + (*n - l[i]);
          z = x + l[i];
          u = x;
          s[i] = 0.0;
          while (u < y)
            s[i] += *u++ * *z++;
        }
    }




```{r}
# unload the library
dyn.unload(paste(filename, .Platform$dynlib.ext, sep = ""))
```
