---
title: "Convolution 1D by Song-Ho"
output: html_notebook
---

http://www.songho.ca/dsp/convolution/convolution.html

The Source command both executes the sourceCpp function and monitors the results of the compilation for errors. If errors do occur then the gcc error log is parsed into a navigable list:

Note that you can also check the Source on Save option to automatically source the C++ file every time it is saved. This variation will keep the keyboard focus within the editor whereas the Source command will move focus to the console.

```{r}
library(Rcpp)
sourceCpp("../src/convolve1D.cpp")

x = c(3, 4, 5, 0, 0)
y = c(2, 1)

convolve1D(x, y)
# 6    11    14     5     0     0
```

```{r}
library(Rcpp)
sourceCpp("../src/convolve1D.cpp")

x = c(3, 4, 5)
y = c(2, 1)

convolve1D(x, y)
# 6    11    14     5
```


```{r}
library(Rcpp)

sourceCpp("../src/convolve1D.cpp")

x <- matrix(rnorm(10))
y <- matrix(rnorm(10))

convolve1D(x, y)
```

http://www.songho.ca/dsp/convolution/convolution.html

```{r}
library(Rcpp)

sourceCpp("../src/convolve2D.cpp")

x <- matrix(1, nrow=3, ncol=4)
y <- matrix(1, nrow=5, ncol=2)

convolve2D(x, y)
```


```{r}
print(x)
print(y)
```

